\subsection{Held-Karp algorithm} \label{algorithm-tsp-heldkarp}
The Held-Karp algorithm is a \gls{dynamic programming} approach to problems similar to the \acrlong*{TSP}, proposed in 1962 independently by Richard Bellman \cite{bellman62} and by Michael Held and Richard Karp \cite{held-karp62}. The main idea is that the current state can be uniquely described by the city $v$ we are at, and the set $S$ of cities we have already visited including $v$, and for each state we know the minimum cost $D(S,v)$ of reaching $v$ after having visited all cities in $S$ including $v$. The algorithm then relies on a trivial recurrent formula (with some additional base cases) to express the links between states in \textsc{HK}. Now that we have all states $(V, u), u \in V \backslash \{s\}$, we pick $u$ so as to minimize $D(V, u)+w(V,u,s)$ (given all nodes were visited, pick the node with the less weight to visit all nodes plus return to the source node $s$).
\begin{algorithm}[H]
    \caption{Held-Karp algorithm}
    \label{alg:held-karp}
    \begin{algorithmic}[1]
        \Function{HK}{$S=0$, $v=s$}
            \If {$D(S,v) \neq \infty$} \Return $D(S,v)$
            \EndIf
            \If {$s \not \in S$} \Return $\infty$ \Comment{$s \not \in S$ is an invalid state}
            \EndIf
            \If {$v = s$}
                \If {$S = \{s\}$} \Return {$0$} \Comment{$D(\{s\}, s)=0$}
                \Else             { \Return {$\infty$} } \Comment{$(S, s)~:~S \neq \{s\}$ is an invalid state}
                \EndIf
            \EndIf
            \For {$u \in S \backslash \{u\}$}
                \State {$c \gets \Call{HK}{S \backslash \{v\}, u} + w(S \backslash \{v\}, u, v)$}
                \If{$c < D(S,v)$}
                    \State {$P(S,v) \gets (S\backslash \{v\}, u)$}
                    \State {$D(S,v) \gets c$} \Comment{$D(S, v)=\min_{u \in S}\{D(S\backslash \{v\}, u) + w(S\backslash\{v\},u,v)\}$}
                \EndIf
            \EndFor
            \State \Return {$D(S,v)$}
        \EndFunction
        \Function{HeldKarp}{$V$, $s$, $w$}
            \For {$S \in \mathscr{P}(V),\,v \in V$} {$D(S,v)=\infty$, $P(S,v)=\text{NULL}$}
            \EndFor
            \State {$c \gets \infty$}
            \For {$u \in V \backslash \{s\}$}
                \State {$c' \gets \Call{HK}{V, u}+w(V,u,s)$}
                \If {$c' < c$}
                    \State {$c \gets c'$}
                    \State {$cur \gets (V, u)$}
                \EndIf
            \EndFor
            \State {$L \gets \langle s \rangle$}
            \Repeat 
                \State {$L \gets \langle cur_u, L \rangle$}
            \Until {$cur_u = s$}
            \State \Return {$L$}
        \EndFunction
    \end{algorithmic}
\end{algorithm}
As we will see, this algorithm has exponential complexity so it is only addequate for very small values of $|V|$, so set $S$ is usually implemented as a bitmask subset of $V$.
\subsubsection{Proof of correctness}
We will rather present an argument as to why these recurrencies are valid. A valid state can either be:
\begin{itemize}
    \item Of the form $(S,v)=(\{s\}, s)$, where the cost is $0$ since the agent has not yet left $s$.
    \item Not of the form $(S,v)=(\{s\}, s)$, in which case $v$ is not the only visited city so far, so among solutions that have visited node $S \backslash \{v\}$ and are at state $u \in S$ we pick the one with the least cost.
\end{itemize}
\subsubsection{Complexity analysis}
There are $2^{|V|}$ subsets of $V$, and $|V|$ possibilities for $v$, meaning there are at most $|V| \cdot 2^{|V|}$ states. Some of these states are not valid or unreachable, but they don't ammount to a value that effectively reduces the time and memory complexities.\par
To find the value of a state $(S,v)$, we will iterate over $|S|$ other states. Since $|S|$ is at most $|V|$, we have a total time complexity of $O(|V|^2 \cdot 2^{|V|})$. The memory complexity for this algorithm is $O(|V| \cdot 2^{|V|})$, since we need to store the result for each state.\par
States of the forms $(S, v)\colon v \not \in S$ are not valid; these ammount to approximately half the states, so we can claim we have approximately $|V|\cdot 2^{|V|-1}$ valid states, and we will have to perform around $|V|^2\cdot 2^{|V|-1}$ operations.\par
\subsubsection{Remarks}
We will have to perform approximately $|V|^2\cdot 2^{|V|-1}$ operations. Using the rule-of-thumb that a computer can perform $10^8$ serial operations per second, and assuming as maximum van capacity the legal limit for european driving license category D1 (light buses with a maximum of 16 seats, not including driver)\cite{dir-2006-126-ec}, we can execute this algorithm in about $\SI{0.08}{s}$, which is acceptable if this algorithm is run only a few times.
