\subsection{Dijkstra's algorithm}
Dijkstra's algorithm is one of the most famous graph algorithms, and the most popular shortest path algorithm between nodes. It was first published in 1959 by Dutch computer scientist Edsger Dijkstra \cite{dijkstra}.\par
This algorithm finds the distance between a node $s$ and every other node $d$ reachable from $s$ (thus solving the shortest path problem in the meantime). The fundamental idea is that, at each moment, we hold a set $Q$ of unprocessed nodes, and we know for all nodes the values of functions distance $dist$ and previous node $prev$ (temporary if the node is in $Q$, final otherwise). We then process the node $u \in Q$ with the least distance to $s$, which consists of iterating over all nodes adjacent to $u$, calculate their potential new distance to $s$ passing through $u$ ($c'=c+w(u,v)$), and if the new distance is smaller we update $dist(v)=c'$ and $prev(v)=u$.
\begin{center}
    \begin{algorithm}[ht]
        \caption{Dijkstra's algorithm}
        \label{alg-dijkstra}
        \begin{minipage}[t]{0.49\linewidth}
            (a) Mathematical version
            \begin{algorithmic}[1]
                \Function{Dijkstra}{$G(V,E)$, $s$}
                    \State $Q \gets \emptyset$
                    \For {$v \in V$}
                        \State {$dist(v) \gets \infty$}
                        \State {$prev(v) \gets \text{NULL}$}
                        \State $Q \gets Q \cup \{v\}$
                    \EndFor
                    \State {$dist(s) \gets 0$}
                    \While {$|Q| > 0$}
                        \State $u \gets \text{element of } Q \text{ with least } dist(u)$
                        \State $Q \gets Q \backslash \{u\}$
                        \For {$v \in \Call{Adj}{G, u}$}
                            \State $c' \gets c + w(u, v)$
                            \If {$c' < dist(v)$}
                                \State $dist(v) \gets c'$
                                \State $prev(v) \gets u$
                            \EndIf
                        \EndFor
                    \EndWhile
                    \State \Return $dist$, $prev$
                \EndFunction
            \end{algorithmic}
        \end{minipage}
        \begin{minipage}[t]{0.49\linewidth}
            (b) Programmatic version
            \begin{algorithmic}[1]
                \Function{Dijkstra}{$Adj$, $s$}
                    \State $Q = \Call{Min\_Priority\_Queue}{ }$
                    \For {$v : V$}
                        \State {$dist[v] = \infty$}
                        \State {$prev[v] = \text{NULL}$}
                        \State {$Q.\Call{Insert\_With\_Key}{\infty, v}$}
                    \EndFor
                    \State $dist[s] = 0$
                    \State $Q.\Call{Decrease\_Key}{s, dist[s]}$
                    \While {$Q.\Call{Size}{ } > 0$}
                        \State $u = Q.\Call{Extract\_Min}{ }$
                        \For {$(v, w): Adj[u]$}
                            \State $c' = c + w$
                            \If {$c' < dist[v]$}
                                \State $dist[v] = c'$
                                \State $prev[v] = u$
                                \State $Q.\Call{Decrease\_Key}{v, dist[v]}$
                            \EndIf
                        \EndFor
                    \EndWhile
                    \State \Return {$dist$, $prev$}
                \EndFunction
            \end{algorithmic}
        \end{minipage}
    \end{algorithm}
\end{center}
For Dijkstra's algorithm we usually represent the graph as an adjacency list $Adj$ (admits iteration in $O(1)$ for each adjacent node). We also use a minimum priority queue for $Q$, where the key of $u$ is $dist(u)$ (fast extraction of element with smallest key and key decrease when processing adjacent nodes).
\subsubsection{Proof of correctness}
The full proof is rather long and is promptly available, thus we will only present a short argument for correctness. The full proof can be found, for instance, in \cite[p.~659]{intro-alg}
\begin{itemize}
    \item \textbf{Pre-conditions:} $G(V, E, w)$ is a graph. $s \in V$. $Q$ contains all nodes. $dist(s)=0$ and $prev(s)=\text{NULL}$
    \item \textbf{Post-conditions:} shortest path distance $dist(u)$ and previous node $prev(u)$ from $s$ to $\forall u \in V$ are final. Distance of unreachable nodes is $\infty$ and previous node in shortest path is invalid. Previous node for $s$ is also invalid.\par
    \item \textbf{Cycle invariant:} $\forall u \not \in Q$, $dist(u)$ is the weight of the shortest path from $s$ to $u$ and $prev(u)$ is the previous node in the shortest path from $s$ to $u$. $\forall u \in Q$, $dist(u)$ is the least distance from $s$ to $u$ going only through nodes not in $Q$.\par
    This holds for the next cycle run because the element $u$ we remove from $Q$ already has the least $dist(u)$ passing through nodes $u' \not \in Q$, and it can't pass through nodes $u' \in Q$ because there is no node with less $dist(u')$ that would allow to reduce $dist(u)$ even further.
    \item \textbf{Cycle variant:} $|Q|$, which is:
    \begin{itemize}
        \item Integer, given a size can only be integer.
        \item Strictly decreasing, given that one node is removed from $Q$ every time the cycle runs.
        \item Positive, given it starts with value $|V|$ where $V$ has at least node $s$. The cycle breaks when $|Q| = 0$.
    \end{itemize}
\end{itemize}
\subsubsection{Complexity analysis}
The initial calls to \textsc{Insert\_With\_Key} are made in time $O(|V|)$, given all keys are $\infty$ and the heap property is maintained. In the main cycle, all elementary operations are done in $O(1)$ except for $Q$ operations \textsc{Extract\_Min} and \textsc{Decrease\_Key}. Let their time complexities be $O(\text{T}_{em})$ and $O(\text{T}_{dk})$ respectively. \textsc{Extract\_Min} is run once for each of the $|V|$ nodes we remove from $Q$. For each node $u$ we iterate over $Adj[u]$, thus iterating once over each of the $|E|$ edges, potentially running \textsc{Decrease\_Key} for each edge. This yields the total time complexity $O(|E|\cdot \text{T}_{dk}+|V|\cdot \text{T}_{em})$. \par
Implementing $Q$ as a binary heap, we have $O(\text{T}_{dk})=O(\log |V|)$ and $O(\text{T}_{em})=O(\log |V|)$, thus yielding the total time complexity $O((|E|+|V|)\log |V|)$. This can be further improved with a Fibonacci heap, with $O(\text{T}_{dk})=O(1)$ and $O(\text{T}_{em})=O(\log |V|)$ (both amortized), thus yielding $O(|E| + |V|\log |V|)$.\par
The algorithm takes as much memory as the heaps, thus giving a memory complexity of $O(|V|)$.
