\subsection{Advanced topics} \label{algorithm-vrp-advanced}
There are various other methods that can be applied to reach decent, and sometimes even close-to-optimal, solutions. Recently there has been a significant effort to study, improve and implement \glspl{metaheuristic} to solve the \acrshort*{VRP}, some of which will be presented, and implement if needed.
\subsubsection{Genetic algorithms}
A genetic algorithm relies in principles of \gls{natural selection} and genetics to find increaingly better solutions to a problem. It is an \gls{iterative} method that consists of applying mutations to a population of solutions (genetical mutation), recombining individual solutions (reproduction) and selecting the most fit individuals (natural selection), until we have iterated over a certain number of generations \cite{intro-genetic-algs}.
\subsubsection{Simulated annealing}
Simulated annealing is an \gls{iterative} process based on thermodynamical principles of cooling. Annealing is a heat treatment that consists of heating a material above its recrystallization temperature, allowing atoms to rearrange and form progressively larger crystal structures towards an equilibrium state of reduced energy.\par
This metaheuristic works in a similar way, where the system has a certain temperature $T$ which is progressively reduced according to an annealing schedule. Let $E_1$ be the energy of the current state. On each step, a new state is determined by randomly changing the current state, its energy $E_2$ is calculated, and a function returns the probability of the system to transition from state 1 to state 2, according to their energies $E_1$, $E_2$ and the temperature $T$, where consecutively lower temperatures must yield lesser probabilities of transitioning to a higher energy state. Thus, there is always a chance, albeit small, that the system may escape from a local minimum and find a better local minimum by going through a higher energy state \cite{numerical-recipes}.
